---
title: "dottable_hw"
author: "Sydney Ball"
date: "`r Sys.Date()`"
output: pdf_document
---
## Libraries

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(data.table)
```

### Load Data

```{r}
nycdata <- fread("~/Desktop/nycdata.csv")  
head(nycdata)
str(nycdata)
```

### Question 1:

Use and show data.table code to select the variables year, month, day, and hour from the imported flights data

```{r}
nycdata[, c("year", "month", "day", "hour"), with = FALSE]
```

### Question 2:

Use and show data. table code to produce a table that shows a carrier of DL, an origin of JFK and a destination of SEA

```{r}
nycdata[carrier =="DL" & origin == "JFK" & dest == "SEA"]
```

### Question 3:

Use and show data.table code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.

```{r}
nycdata[carrier == "UA" & month == 3 & air_time < 330]
```

### Question 4:

Use and show tidyverse code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.

```{r}
nycdata |> 
  filter(carrier == "UA", month == 3, air_time < 330)
```

### Question 5:

Use the data.table method to add a variable called speed that is the average air speed of the plane in miles per hour.

```{r}
nycdata[, speed := distance / (air_time / 60)]
head(nycdata[, .(distance, air_time, speed)])
```

### Question 6:

Use the tidyverse method to add a variable called speed that is the average air speed of the plane in miles per hour.

```{r}
nycdata |>
  mutate(speed = distance / (air_time / 60))
```

### Question 7:

Show and use coding to change the carrier abbreviation of UA to UniitedAir, 7a. data.table method 7b. tidyverse method (Use a sequence of dplyr commands so that you can see the change in your table

```{r}
#7a
nycdata[carrier == "UA", carrier := "UnitedAir"]
head(nycdata[, .(carrier = "UnitedAir", origin, dest)])
```

```{r}
#7b
nycdata |>
  mutate(carrier = ifelse(carrier == "UA", "UnitedAir", carrier)) |>
  filter(carrier == "UnitedAir")
```